---
title: "Assignment 1 - Language Development in ASD - part 4"
author: "Alberte Seeberg"
date: "August 10, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Welcome to the fourth exciting part of the Language Development in ASD exercise

In this exercise we will assess how many participants we would need to adequately replicate our findings (ensuring our sample size is adequate, our alpha at 0.05 and our beta at 0.8).

### Exercise 1

How much power does your study have (if your model estimates are quite right)?
- [GitHub]Load your dataset, fit your favorite model, assess power for your main effects and interactions of interest.
- Report the power analysis and comment on what you can (or cannot) use its estimates for.

```{r}
#set working directory
setwd("~/Desktop/3. Semester/Experimental methods 3/Portfolio 1/Portfolio 1")

#load packages
library(MuMIn)
library(lmerTest)
library(modelr)
library(plyr)
library(stringr)
library(tidyverse)
library(merTools)
library(caret)
library(simr)
library(MASS)
library(dplyr)

#Read dataset
ClData = read.csv("clean_data.csv")
ClData = ClData[-1]

#Making my model
model1=lmer(CHI_MLU~1+VISIT+Diagnosis+(1+VISIT|ID), ClData, REML=FALSE)
plot(model1)
summary(model1)

#Getting the power for my model with visits as fixed effect
PowerV=powerSim(model1,fixed("VISIT"),nsim=100)
PowerV

#Making my model with an interaction effect
model2=lmer(CHI_MLU~1+VISIT*Diagnosis+(1+VISIT|ID), ClData, REML=FALSE)
plot(model2)
summary(model2)

PowerV2=powerSim(model1,fixed("Diagnosis"), nsim=100)
PowerV2

PowerV3=powerSim(model2, fixed("VISIT:Diagnosis"), nsim=100)
PowerV3


```

### Exercise 2

How would you perform a more conservative power analysis?
- Identify and justify a minimum effect size for each of your relevant effects
- [GitHub] take the model from exercise 1 and replace the effects with the minimum effect size that you'd accept.
- [GitHub] assess the power curve by Child.ID, identifying an ideal number of participants to estimate each effect
- OPTIONAL if your power estimates do not reach an acceptable threshold simulate additional participants and repeat the previous analysis
- Report the power analysis and comment on what you can (or cannot) use its estimates for.

```{r}

fixef(model1)["VISIT"] = 0.05
fixef(model1)["DiagnosisTD"] = -0.1
fixef(model2)["VISIT:DiagnosisTD"] = 0.15
fixef(model2)["VISIT"] = 0.5
fixef(model2)["DiagnosisTD"] = -0.1
powerCurveV = powerCurve(model1,fixed("VISIT"),along="ID",nsim=100)
plot(powerCurveV)
powerCurveV
powerCurveV2 = powerCurve(model1,fixed("Diagnosis"),along="ID",nsim=100)
plot(powerCurveV2)
powerCurveV2
powerCurveV3 = powerCurve(model2,fixed("VISIT:Diagnosis"),along="ID",nsim=100)
plot(powerCurveV3)
powerCurveV3
model1Normalbeta = lmerTest::lmer(CHI_MLU ~ 1 + VISIT*Diagnosis + (1 + VISIT|ID), ClData, REML = FALSE)

### Riccardo's clumsy function to simulate new participants
### TO DO points are only notes for myself, so not part of the assignment

createNewData <- function (participants,visits,model){
  # participants is the number of subjects
  # visits is the number of visits
  # TO DO: LOOP THROUGH ALL FE ROWS AND AUTOMATICALLY EXTRACT NAMES OF FIXED EFFECTS AND ESTIMATES
  fe <- fixef(model)
  Intercept <- fe[1] #intercept
  bVisit <- fe[2] #visit
  bDiagnosis <- fe[3] #diagnosis
  bVisitDiagnosis <- fe[4] #visit diagnosis interaction
  # TO DO: INTEGRATE STANDARD ERROR?
  
  # TO DO: LOOP THROUGH ALL VC COMPONENTS AND AUTOMATICALLY EXTRACT NAMES OF EFFECTS AND ESTIMATES
  vc<-VarCorr(model) # variance component
  sigmaSubject <- as.numeric(attr(vc[[1]],"stddev")[1]) # random intercept by subject
  sigmaVisit <- as.numeric(attr(vc[[1]],"stddev")[2]) # random slope of visit over subject
  sigmaResiduals <- as.numeric(attr(vc,"sc"))
  sigmaCorrelation <- as.numeric(attr(vc[[1]],"correlation")[2])
  
  # Create an empty dataframe
  d=expand.grid(VISIT=1:visits,ID=1:participants)
  # Randomly sample from a binomial (to generate the diagnosis)
  condition <- sample(rep(0:1, participants/2))
  d$Diagnosis<-condition[d$ID]
  d$Diagnosis[is.na(d$Diagnosis)]<-1
  
  ## Define variance covariance matrices:
  Sigma.u<-matrix(c(sigmaSubject^2,
                    sigmaCorrelation*sigmaSubject*sigmaVisit,
                    sigmaCorrelation*sigmaSubject*sigmaVisit,
                    sigmaVisit^2),nrow=2)
  
  ## generate new fake participants (column1=RandomIntercept, column2=RandomSlope)
  u<-mvrnorm(n=participants,
             mu=c(0,0),Sigma=cov(ranef(model)$ID))
  
  ## now generate fake data:
  ### the outcome is extracted from a gaussian with
  ### the solution to the model's equation as mean and
  ### the residual standard deviation as standard deviation 
  d$CHI_MLU <- rnorm(participants*visits,
                     (Intercept+u[,1]) +
                     (bVisit+u[,2])*d$Visit + 
                     bDiagnosis*d$Diagnosis ,sigmaResiduals)  
  
  return(d)
}

# To simulate additional participants 
fakedata = createNewData(100, 6, model1Normalbeta)

# To create new IDs
fakedata$ID=fakedata$ID + 100
fakedata$Diagnosis=as.factor(fakedata$Diagnosis)
fakedata$Diagnosis=revalue(fakedata$Diagnosis, c("0"="ASD", "1"="TD"))


# To select relevant variables
DataSimple=dplyr::select(ClData, ID, VISIT, Diagnosis, CHI_MLU)

# To bind together data
newdata=rbind(fakedata, DataSimple)

# To create new model
Model3 = lmerTest::lmer(CHI_MLU ~ 1 + VISIT+Diagnosis + (1 + VISIT|ID), newdata, REML = FALSE)
summary(Model3)

# To set minimum interesting effect sizes
fixef(Model3)["VISIT"] = 0.05
fixef(Model3)["DiagnosisTD"] = -0.10

powerCurveD2 = powerCurve(Model3, fixed("Diagnosis"),along="ID", nsim=10)
plot(powerCurveD2)
powerCurveD2

powerCurveV2 = powerCurve(Model3, fixed("VISIT"),along="ID", nsim=10)
plot(powerCurveV2)
powerCurveV2

```


### Exercise 3

Assume you have only the resources to collect 30 kids (15 with ASD and 15 TDs). Identify the power for each relevant effect and discuss whether it's worth to run the study and why.


